Predicting Solid State Material Platforms for Quantum Technologies

### 2021-12-11 paper.tex
Updates from Overleaf

%We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25000$ materials. Furthermore, we have developed and implemented three data mining approaches, termed \textit{the Ferrenti approach}, \textit{the augmented Ferrenti approach} and \textit{the insightful approach} for defining three distinct training sets for the supervised machine learning algorithms logistic regression, decision tree, random forest and gradient boost to be trained on.
Data extraction tools have been developed for a number of materials science databases resulting in a data set consisting of over $25'000$ materials. The information from the databases was converted into nearly $5'000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data mining process: \textit{The Ferrenti approach}, \textit{the augmented Ferrenti approach}, and \textit{the intuitive approach}. 
%Rename insightful approach! To what? 
%Options include stringent, strict, cherry-picked, informed and intuitive. 
The three approaches define distinct training sets for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

%We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set, whereas the restrictions set in the insightful approach proved suitable. All models for the insightful approach agreed on $214$ predicted candidates, with examples such as ZnGeP$_2$, MgSe, CdS, BP, BC2N, BP, Ge, GeC, InP, and InAs. All three approaches and all ML models agreed on a subset of $47$ eligible candidates of $8$ elemental, $29$ binary, and $10$ tertiary compounds.
We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set. The restrictions set in the intuitive approach, on the other hand, which relies exclusively on findings from the literature, proved more suitable.  
All four machine learning (ML) methods applied to the data set derived using the intuitive approach agreed on a set of $214$ predicted candidate materials to a probability of $>50 \ \%$. 
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates (to a probability of $>XX \ \%$) of $8$ elemental, $29$ binary, and $10$ tertiary compounds.

@@

\section*{Introduction}
The second quantum revolution is heralding the development of new solutions and devices based on principles from quantum mechanics \cite{Acin2018}. Among the promising technologies that are already available we find in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum computer architectures and prototypes. 
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Simultaneously, the concepts of entanglement and teleportation may enable advanced quantum communication protocols such as quantum cryptography and the quantum internet.  
Recent advances have highlighted the potential of quantum computer architectures to outperform classical systems as a 53-qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a 200'000 core supercomputer \cite{Arute_2019}, spurring further investigations into competitive quantum technology platforms. Most recently, IBM announced its 127 qubit quantum processor~\cite{IBM2021}.

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
%    \includegraphics[width=0.9\textwidth]{figures/ht-workflow.jpg}
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new.png}
    \caption{[PLACEHOLDER] Schematic of an example workflow in material informatics. 
    %\marianne{Sebastian makes simplified figure.}
    }
    \label{fig:ht-workflow}
\end{figure}

@@

The majority of discoveries of quantum emitters in semiconductors has so far happened by serendipity, and there is an urgent need for 
a better understanding of how the QT compatible materials properties manifest. 
In this context, a framework for dedicated materials search and analysis is needed. 
The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. The fourth paradigm instead introduces the potential of targeted search for promising material systems to act as hosts for qubit and single-photon emitter defects.  
Instead of searching through a host of signals for the one that matches our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies. 
This is made feasible by the availability of databases containing materials properties for a wide range of different systems. In this work, the data in question is provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of qubit candidates. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

@@

The information stream in this work can be regarded as many modular parts connected in logical pieces. The initial step for gathering material data and building features is visualized by the the flowchart in Figure~\ref{fig:flowchart-makedata}. 
Initially, we start by extracting all entries in
the Materials Project (MP) that matches a specific query. 
The data in MP is calculated using density functional theory (DFT) as implemented in the Vienna Ab Initio Package (VASP) \cite{Kresse1996} and using the The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional to calculate the electronic structure of the materials.

@@

%Marianne: I got some stuff from the thesis 
Before applying any machine learning algorithm, the data was transformed into a numerical representation that reflects the relationship between the input and output data. This transformation is known as generating descriptors
or features, however, we will in this work adapt the name featurization.
The open-source library of Matminer provides many tools to featurize existing features extracted from the Materials Project.
A schematic describing the  featurization process in Matminer is shown in Figure~\ref{fig:flowchart-featurization} 
and focus on a material's composition, structure, atomic sites, density of states and band structure. 
%\textbf{[Describe it and say something about which features we choose.]}  
The 39 features selected in this work are summarized in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not precisely know which features best describe a suitable  host material.

@@

After compiling a data set from the procedure outlined above, we face an obstacle in terms of defining a training set for the machine learning algorithms. This is challenging for two reasons. 
Firstly, although defects like the NV center in diamond and emitters in SiC are becoming
increasingly well studied, few candidate materials are known. An additional consideration
is that the physical mechanisms promoting favorable defect properties are not fully understood. Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
%This is not only challenging due to the lack of known suitable candidates (1),  but also due to the intricacy of defining materials as unsuitable candidates (0). 
We follow three separate procedures to define a training set consisting of materials that are labeled as both suitable and unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to defining a training set is based on the criteria from the paper by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps. 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero  isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Ru, Os, Fe or Ni because of their paramagnetism.
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. 
    \item Ensure that energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

@@

The following steps constitute the process of choosing \textbf{suitable} candidates 
in the \emph{augmented Ferrenti approach}. 
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of zero spin isotopes, except Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV as calculated by PBE-GGA. The higher band gap requirement is included here to avoid including an unfeasible number of candidate materials. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

@@

The suitable candidates in the intuitive approach contain only compounds that are either elementary (unary) or binary. We do not want to discriminate based on the number of elements in a compound, therefore we remove the feature that describes the number of elements in a compound. After the three stages, a total of $187$ entries were labeled as suitable candidates.
See the Supplementary Material at \cite{supplementary} for a table containing the manually screened materials that were labeled as suitable candidates in the training set for the insightful approach.

Since the data set constituting suitable candidates in the third approach contains few entries, we add $400$ \emph{unsuitable} candidates. These are picked at random from the pool of unsuitable candidates from the two previous approaches, in addition to those that were marked as unsuitable through the manual screening process.

\subsubsection*{Comparing the approaches}
The three different methods for data mining place a particular emphasis on each of their particular goals. The Ferrenti approach depends on choosing only elements with spin zero isotopes together with practical filters, while the augmented Ferrenti approach allows a larger variety of elements and removes the practical reasons for excluding candidate materials. Thus, the first approach targets a more narrow prediction space than the second approach does. However, perhaps the most restricted approach is the insightful approach. Since the variety of known suitable materials is substantially more restricted than for the two other approaches, we would expect the insightful approach to provide an even narrower prediction space.

%\marianne{Short paragraph here (or somewhere else) highlighting the potential issues with the two first approaches in terms of actually identifying any trends.}

We provide a visualization of each approach's training data as a parallel coordinate plot for a few selected features in Figure~\ref{fig:parallel-coordinates-approaches}. Parallel coordinate schemes \cite{Inselberga1990, Inselberg1985} represent a multi-dimensional data tuple as one polyline crossing a parallel axis. The selected features are found on the x-axis, while the y-axis shows the value of the data present. Thus, parallel coordinate plots can turn many-dimensional data into a compact two-dimensional representation. Due to possible data cluttering, the figure visualizes a random sample of each class with an upper limit of $250$ per class with transparent lines. The green and red polylines represent suitable and unsuitable candidates, respectively.

The Ferrenti approach and the augmented Ferrenti approach share several similarities; 
for example, all unsuitable candidates exhibit polar space groups (follows from the definition of the data sets), and their upper limits for both ionic character and covalent range are equal for suitable candidates.  
%such as having only unsuitable candidates with polar space groups and having an equal amount of upper limit for both ionic character and covalent range for suitable characters. 
Additionally, the two approaches share that suitable candidates can consist of up to five different elements. 
Interestingly, we observe that
the entries in Fig.~\ref{fig:parallel-coordinates-approaches} span roughly the same dimensions for the Ferrenti and augmented Ferrenti approaches, despite the more restricted nature of the former. 
%even if the augmented Ferrenti approach is less restricted, it appears that the entries span the same dimensions.

The biggest difference is found for the insightful approach. The entries labeled as suitable candidates do not exhibit any magnetization, even though the materials crystallize in both polar and non-polar space groups. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller space in the case of the insightful approach than for the Ferrenti-based approaches. 
Here, we can begin to deduce some trends, if the findings from the insightful approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range appear more important. 
Identification of the features that promote quantum compatible characteristics could be crucial for targeted quantum device design based on defect engineering.

@@

To visualize the complexity of the training sets, we have found the two largest eigenvalues of the covariance matrix of the initial data from the Materials Project, and transformed the training sets according to the corresponding two eigenvectors. The resulting scatter plots are found in Figure~\ref{fig:2dscatterplotpca}. 
Green squares display suitable candidates and red triangles represent unsuitable candidates. 
Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the augmented Ferrenti approaches are largely overlapping. 
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
%Nonetheless, we note that the unsuitable candidates appear to be restricted to a certain parameter space while the suitable candidates extend much further.  
Therefore, we can expect that the two Ferrenti approaches would need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it could prove challenging to find a generalized model based on the available data sets.

In the case of the intuitive approach, on the other hand, we can already start to see a trend where the upper left part of the figure is dominated by suitable candidates. The unsuitable candidates are similarly restricted to the lower right corner and there appears to be a clear barrier between the two, albeit with some exceptions. This indicates that the intuitive approach may be capable of separating materials based on their underlying properties.

@@

It seems likely that the differences in Fig.~\ref{fig:2dscatterplotpca} can be attributed to the specific choices made when extracting data in each of the three approaches. Where the insightful approach takes advantage of the experimental knowledge that is already available in the literature, the Ferrenti and augmented Ferrenti schemes 
%procedures 
rely on previous assumptions for which material properties are responsible for the desired quantum effects.   
An additional factor to consider is the use of practical considerations when restricting the parameter space for material choice. 
In short, a much clearer separation is observed for the intuitive approach as compared to the others. 
These observations constitute an important finding of the work, as we show that it is possible to distinguish between suitable and unsuitable material candidates for QT. 
The manual approach for selecting suitable candidates based on experimental findings yields strong indication that a trend does exist and thus is possible to identify. 
Importantly, the presented methodology for selecting candidates as presented in the intuitive approach can be adapted also to other material systems where a specific set of properties are desired (and observed or predicted for a sufficient number of materials) but the specific mechanisms behind them are unknown.

It is important to note that the constructed data sets encompass compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures. We note that this complexity is not necessarily reflected in the descriptors. 
%Additionally, we acknowledge that many compositions deemed as suitable candidates consist of either rare or dangerous elements. By utilizing an enormously large database as Materials Project, we have to account for their ultimate goal - to model all possible materials and their properties.
%The automated process of adding an entry to their database does not necessarily contain all relevant information about a respective material. This is information that needs to be added manually.
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has happened in the first part of the process, the error follows the entire process and will get increasingly harder to detect. Therefore, we are dependent on that the Materials Project has obtained data with high quality, and we note that it is likely that there are some errors present in our data sets.

Another potential source of uncertainty which should be mentioned is the use of the PBE functional for computing the information in the databases employed herein. Despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput.

\subsection*{Predicting Suitable Material Hosts for Quantum Technology}

Taking the four machine learning algorithms outlined above and applying them to the data sets optimized using each of the three approaches yields twelve sets of results for predicting  materials as suitable material hosts for QT. In this section, we present representative ML results based on each of the three data mining approaches.

\subsubsection*{The Ferrenti Approach}
%Should we include materials in training set? Or keep it, but move into data mining discussion?
We first consider the machine learning classification of the test set based on the Ferrenti approach. 
Notably we find that carbon in a diamond-like structure is classified as a suitable candidate from the training set in the Ferrenti approach. 
This is in agreement with the suitable candidates that were defined manually for the insightful approach. 
Interestingly, from the training set of the Ferrenti approach, we also find that carbon in two-dimensional graphite-like structures is labeled as suitable. 
%Out of the known suitable candidates defined for the insightful approach, we find carbon in diamond-like structures, but we also find two-dimensional carbon in graphite-like structures labeled as suitable. 
All structures of silicon are also labeled as suitable together with one entry of silicon carbide. 
Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present.

%\input{figures/tikz-plots/ferrenti-approach.tex}
%\input{figures/tikz-plots/augmented-ferrenti-approach.tex} 
%\input{figures/tikz-plots/insightful-approach.tex}

@@

The number of candidates predicted by each of the four machine learning methods 'Logistic regression', 'Decision trees', 'Random forests' and 'Gradient boosting' is labeled in Figure~\ref{fig:predictions}(a) for the Ferrenti approach. 
From the training set of $23623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates. All four methods agree on $6804$ suitable candidates, however, many of the materials are predicted with the probability similar to a coin flip. 
If we were to raise the minimum bar of a prediction to
$0.7$ instead of $0.5$, the above ML approaches would only agree on around $3000$ suitable candidates. 
Of the known suitable materials that were present in the training set, we find that all four methods admit almost all materials with a chemical formula matching the known candidates. This can allow materials with unfortunate structures to be labeled as suitable candidates by all of our ML methods. Consequently, the ML methods do not recognize the strict band gap restriction, which makes it challenging to facilitate deep defects. Ideally, we would expect that the ML methods would have probabilities lower than $0.5$ for all materials with band gaps lower than $0.5$~eV, which would be expected behavior based on the training set, but this is not the case. We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all ML methods studied here.

\subsubsection*{The augmented Ferrenti Approach}

>>>>

%We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25000$ materials. Furthermore, we have developed and implemented three data mining approaches, termed \textit{the Ferrenti approach}, \textit{the augmented Ferrenti approach} and \textit{the intuitive approach} for defining three distinct training sets for the supervised machine learning algorithms logistic regression, decision tree, random forest and gradient boost to be trained on.
Data extraction tools have been developed for a number of materials science databases resulting in a data set consisting of over $25.000$ materials. The information from the databases was converted into nearly $5000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data mining process: \textit{The Ferrenti approach}, \textit{the augmented Ferrenti approach}, and \textit{the intuitive approach}. 
%Rename intuitive approach! To what? 
%Options include stringent, strict, cherry-picked, informed and intuitive. 
The three approaches define distinct training sets for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

%We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set, whereas the restrictions set in the intuitive approach proved suitable. All models for the intuitive approach agreed on $214$ predicted candidates, with examples such as ZnGeP$_2$, MgSe, CdS, BP, BC2N, BP, Ge, GeC, InP, and InAs. All three approaches and all ML models agreed on a subset of $47$ eligible candidates of $8$ elemental, $29$ binary, and $10$ tertiary compounds.
We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set. The restrictions set in the intuitive approach, on the other hand, which relies exclusively on findings from the literature, proved more suitable.  
All four machine learning (ML) methods applied to the data set derived using the intuitive approach agreed on a set of $214$ predicted candidate materials to a probability of $>50 \ \%$. 
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates (to a probability of $>XX \ \%$) of $8$ elemental, $29$ binary, and $10$ tertiary compounds.

@@

\section*{Introduction}
The second quantum revolution is heralding the development of new solutions and devices based on principles from quantum mechanics \cite{Acin2018}. Among the promising technologies that are already available we find in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum computer architectures and prototypes. 
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Simultaneously, the concepts of entanglement and teleportation may enable advanced quantum communication protocols such as quantum cryptography and the quantum internet.  
Recent advances have highlighted the potential of quantum computer architectures to outperform classical systems as a 53-qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a $200.000$ core supercomputer \cite{Arute_2019}, spurring further investigations into competitive quantum technology platforms. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}.

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new.png}
    \caption{Schematic of an example workflow in material informatics: results from theory, experiment and computation are fed into material databases. A cycle involving material screening, machine learning and predictions leads to knowledge gain and ultimately applications in fields such as clean energy and quantum technology. 
    }
    \label{fig:ht-workflow}
\end{figure}

@@

The majority of discoveries of quantum emitters in semiconductors has so far happened by serendipity, and there is an urgent need for 
a better understanding of how the QT compatible materials properties manifest. 
In this context, a framework for dedicated materials search and analysis is needed. 
The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. The fourth paradigm instead introduces the potential of targeted search for promising material systems to act as hosts for qubit and single-photon emitter defects.  
Instead of searching through a host of signals for the one that matches our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies, following the framework illustrated in Figure~\ref{fig:ht-workflow}. 
This is made feasible by the availability of databases containing materials properties for a wide range of different systems. In this work, the data in question is provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of qubit candidates. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

@@

The information stream in this work can be regarded as many modular parts connected in logical pieces. The initial step for gathering material data and building features is visualized by the the flowchart in Figure~\ref{fig:flowchart-makedata}. 
Initially, we start by extracting all entries in
the Materials Project (MP) that matches a specific query. 
The MP database contains ground state properties of different materials that are calculated using density functional theory (DFT) as implemented in the Vienna Ab Initio Package (VASP) \cite{Kresse1996} and using the The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional to calculate the electronic structure of the materials.

@@

Before applying any machine learning algorithm, the data was transformed into a numerical representation that reflects the relationship between the input and output data. This transformation is known as generating descriptors
or features, however, we will in this work adapt the name featurization.
The open-source library of Matminer provides many tools to featurize existing features extracted from the Materials Project.
A schematic describing the  featurization process in Matminer is shown in Figure~\ref{fig:flowchart-featurization} 
and focus on a material's composition, structure, atomic sites, density of states and band structure. 
%\textbf{[Describe it and say something about which features we choose.]}  
The 39 features selected in this work are summarized in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not precisely know which features best describe a suitable  host material.

@@

After compiling a data set from the procedure outlined above, we face an obstacle in terms of defining a training set for the machine learning algorithms. This is challenging for two reasons. 
Firstly, although defects like the NV center in diamond and emitters in SiC are becoming
increasingly well studied, few candidate materials are known. An additional consideration
is that the physical mechanisms promoting favorable defect properties are not fully understood. Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
We follow three separate procedures to define a training set consisting of materials that are labeled as both suitable and unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to defining a training set is based on the criteria from the paper by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps. 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero  isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Ru, Os, Fe or Ni because of their paramagnetism.
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. The value of $0.5$~eV was chosen to match that typically predicted for silicon by PBE-level DFT calculations. 
    \item Ensure that energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

@@

The following steps constitute the process of choosing \textbf{suitable} candidates 
in the \emph{augmented Ferrenti approach}. 
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of zero spin isotopes, except Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV as calculated by PBE-GGA. The higher band gap requirement (as compared to the Ferrenti approach) is included here to avoid including an unfeasible number of candidate materials. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

@@

The suitable candidates in the intuitive approach contain only compounds that are either elementary (unary) or binary. We do not want to discriminate based on the number of elements in a compound, therefore we remove the feature that describes the number of elements in a compound. After the three stages, a total of $187$ entries were labeled as suitable candidates.
See the Supplementary Material at \cite{supplementary} for a table containing the manually screened materials that were labeled as suitable candidates in the training set for the intuitive approach.

Since the data set constituting suitable candidates in the third approach contains few entries, we add $400$ candidates that are labeled as \emph{unsuitable}. These are picked at random from the pool of unsuitable candidates from the two previous approaches, in addition to those that were marked as unsuitable through the manual screening process.

\subsubsection*{Comparing the approaches}
The three different approached to data mining place a particular emphasis on each of their particular goals. The Ferrenti approach depends on choosing only elements with spin zero isotopes together with practical filters, while the augmented Ferrenti approach allows a larger variety of elements and removes the practical reasons for excluding candidate materials. Thus, the first approach targets a more narrow prediction space than the second approach does. However, perhaps the most restricted approach is the intuitive one. Since the variety of known suitable materials is substantially more restricted than for the two other approaches, we would expect the intuitive approach to provide an even narrower prediction space. Here, one should keep in mind that ...

We provide a visualization of each approach's training data as a parallel coordinate plot for a few selected features in Figure~\ref{fig:parallel-coordinates-approaches}. Parallel coordinate schemes \cite{Inselberga1990, Inselberg1985} represent a multi-dimensional data tuple as one polyline crossing a parallel axis. The selected features are found on the x-axis, while the y-axis shows the value of the data present. Thus, parallel coordinate plots can turn many-dimensional data into a compact two-dimensional representation. Due to possible data cluttering, the figure visualizes a random sample of each class (suitable or unsuitable) with an upper limit of $250$ per class with transparent lines. The green and red polylines represent suitable and unsuitable candidates, respectively.

The Ferrenti approach and the augmented Ferrenti approach share several similarities. 
Interestingly, we observe that the entries in Fig.~\ref{fig:parallel-coordinates-approaches} span roughly the same dimensions for the Ferrenti and augmented Ferrenti approaches, despite the more restricted nature of the former.  
The two approaches' upper limits for ionic character and covalent range are very similar in the case of candidates labeled as both suitable and unsuitable.   
Moreover, all candidates that were labeled unsuitable exhibit polar space groups (follows from the definition of the data sets), and the two approaches also share that suitable candidates can consist of up to five different elements.

The biggest difference is found for the intuitive approach. The entries labeled as suitable candidates do not exhibit any magnetization, even though the materials crystallize in both polar and non-polar space groups - contrarily to the two Ferrenti-based approaches. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller space in the case of the intuitive approach than for the two former ones.  
Here, we can begin to deduce some trends, if the findings from the intuitive approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range appear more important. 
Identification of the features that promote quantum compatible characteristics could be crucial for targeted quantum device design based on defect engineering.

@@

To visualize the complexity of the training sets, we have found the two largest eigenvalues of the covariance matrix of the initial data from the Materials Project, and transformed the training sets according to the corresponding two eigenvectors. The resulting scatter plots are found in Figure~\ref{fig:2dscatterplotpca}. 
Green squares display suitable candidates and red triangles represent unsuitable candidates. 
Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the augmented Ferrenti approaches are largely overlapping. 
The logic behind categorizing materials in the two classes (suitable and unsuitable) appears to not translate into a distinct separation, at least not in the representation of Fig.~\ref{fig:2dscatterplotpca}. 
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
%Nonetheless, we note that the unsuitable candidates appear to be restricted to a certain parameter space while the suitable candidates extend much further.  
Therefore, we can expect that the two Ferrenti approaches would need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it could prove challenging to find a generalized model based on the available data sets.

In the case of the intuitive approach, on the other hand, we can already start to see a trend where the upper left part of the figure is dominated by suitable candidates. The unsuitable candidates are similarly restricted to the lower right corner and there appears to be a clear barrier between the two, albeit with some exceptions. This indicates that the intuitive approach may be capable of separating materials based on their underlying properties.  
Thus, devising a sound framework for the data mining process is paramount to obtain a suitable training set for the machine learning methods.

@@

It seems likely that the differences in Fig.~\ref{fig:2dscatterplotpca} can be attributed to the specific choices made when extracting data in each of the three approaches. Where the intuitive approach takes advantage of the experimental knowledge that is available in the literature, the Ferrenti and augmented Ferrenti schemes rely on assumptions for which material properties are responsible for the desired quantum effects.   
An additional factor to consider is the use of practical considerations (e.g., toxic materials or the availability of spin zero isotopes) when restricting the parameter space for material choice. 
Intriguingly, a much clearer separation is observed for the intuitive approach, where no restrictions were imposed in terms of specific characteristics for the materials, as compared to the two others. 
These observations constitute an important finding of the work, as we show that it is possible to distinguish between suitable and unsuitable material candidates for QT. The manual approach for selecting suitable candidates based on experimental findings yields strong indications that a trend does exist and thus is possible to identify. 
Importantly, the methodology for selecting candidates as implemented in the intuitive approach can be adapted also to other material systems where a specific set of properties is desired (and observed or predicted for a sufficient number of materials) but the specific mechanisms behind them are unknown.

It is important to note that the constructed data sets encompass compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures. We note that this complexity is not necessarily reflected in the descriptors. 
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has happened in the first part of the process, the error follows the entire process and will get increasingly harder to detect. Therefore, we are dependent on that the Materials Project has obtained data with high quality, and we note that it is likely that there are some errors present in our data sets.

Another potential source of uncertainty that should be mentioned is the use of the PBE functional for computing the information in the databases employed herein. Despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor properties.

\subsection*{Predicting Suitable Material Hosts for Quantum Technology}
Four well-tested machine learning algorithms were employed in this work: logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{Mehta2019,Hastie2009}. 
These four machine learning methods were applied to the data sets that were optimized using each of the three approaches (Ferrenti, augmented Ferrenti and intuitive) outlined above. This yielded twelve sets of materials that were predicted as suitable material hosts for QT applications. In this section, we present representative results after applying machine learning methods to the datasets resulting from each of the three data mining approaches.

\subsubsection*{The Ferrenti approach}
We first consider the machine learning classification of the test set derived in the Ferrenti approach. 
Taking a closer look at the training set from the Ferrenti approach, we find that carbon in a diamond-like structure is classified as a suitable candidate. 
This is in good agreement with experimental observations (see Table~\ref{tab:qt-materials}) and hence also the candidates that were manually defined as suitable in the intuitive approach.  
Interestingly, in the training set from the Ferrenti approach, we also find that carbon in two-dimensional graphite-like structures are labeled as suitable. 
All structures of silicon are also labeled as suitable together with one entry of silicon carbide. Note that this is the 3C polytype of SiC, meaning that the most quantum compatible SiC polytype, 4H-SiC, is missing from the training set. Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present in the training data.

@@

The number of candidates predicted by each of the four machine learning methods 'Logistic regression', 'Decision trees', 'Random forests' and 'Gradient boosting' is labeled in Figure~\ref{fig:predictions}(a) for the Ferrenti approach. 
From the training set of $23.623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates. All four methods agree on a total of $6804$ suitable candidates, however, many of these materials are predicted with the probability similar to a coin flip. 
If we were to raise the minimum bar of a prediction to $70 \ \%$ instead of $50 \ \%$, the above ML algorithms would only agree on around $3000$ suitable candidates. 
We find that all four ML methods admit almost all materials with a chemical formula matching the known suitable candidates (see Table~\ref{tab:qt-materials}) that were present in the training set. 
This can allow materials with unfortunate structures to be labeled as suitable candidates by all of our ML methods. 
%Be careful here - PBE is known for underestimating band gaps. 
For example, the ML methods do not maintain the band gap restriction from the training set definition. 
Ideally, we would expect that the ML methods would have probabilities lower than $50 \ \%$ for all materials with band gaps lower than $0.5$~eV, which would be expected behavior based on the training set, but this is not the case. 
We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all four ML methods employed herein, 
indicating that the band gap restriction is not recognized as important by the machine learning algorithms. 
Small band gaps could make it challenging to facilitate deep level defects. 
On the other hand, due to the known underestimation of band gaps by the PBE functional, the band gaps could in reality be larger for some of the materials.
